### 2011-02-12 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
changes suggested by Alex are done, and I have added new sections. not yet finished!
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1265 d8b0c59b-6544-0410-94d9-83a5658a058e

\usepackage{inputenc}
\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsmath}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{6.6cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
 & & Ismael Rivera (NUIG) & \\ \hline
2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Second version of the deliverable \\
 & & Ismael Rivera (NUIG) & \\ \hline
\end{tabular}
\end{small}

@@

changes from M24 version:
\begin{itemize}
 \item 'find' and 'check' operations for screens (screen-flow design) has been removed. the same functionaly was offered by 'find\&check' doing just 1 request.
 \item API calls (both find\&check, planner, metadata) updated to reflect the changes in the prototype
 \item Adds operation for building block cloning.
 \item added appendix explaining how the recommendation/planning is done
 \item architecture have been modified reflecting the changes in the catalogue's architecture, and including new components such as the machine learning library (Mahout) and a library to query iServe.
 \item uses Sindice in the 'ontology importer' mechanism
\end{itemize}

@@

The following section covers the configuration options you may have to set up before start using the Catalogue. The FAST Catalogue relies on a Sesame repository for the persistent storage. Sesame repositories can be local or remote. Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat. The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint. These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.

Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed. If you are testing the application in your machine, it may be something like \verb|http://catalogueURL/FASTCatalogue|.

Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory. For instance:

@@

Nevertheless, setting up a remote repository requires a few more steps. First the OpenRDF Sesame Server needs to be installed. The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}. Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier. For instance:

\begin{verbatim}
sesameServer=http://localhost:8880/openrdf-sesame
repositoryID=repository-ex1
\end{verbatim}

@@

It is recommended to check if the FAST Catalogue has been successfully installed. To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of building block such as Screen. The URL would be similar to \verb|http://catalogueURL/FASTCatalogue/screens|. Depending on the variant chosen, the response may be different. For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.

@@

This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or APIs is done. The Catalogue platform has a modular design, which are spread in basically three layers, depending on the nature of each component:
\begin{itemize}
 \item The top layer represents the public API for the Catalogue. It puts together all the available information and services to offer the maximum powerful and flexibility for other IDEs or applications. In FAST, here is where the Gadget Visual Storyboard
 \item The core layer implements the object model, and provides the business logic for the different processes involved in the Catalogue, integrating third party libraries where needed.
 \item The pumbling layer offers an abstraction to RDF model logic and triple stores.
\end{itemize}

\begin{figure}[htb]
\label{fig:catalogue_architecture}
\begin{center}
	\includegraphics[width=12cm]{images/catalogue_architecture}
	\caption{Catalogue's Architecture}
\end{center}
\end{figure}

\subsection{Presentation layer / API}

The Presentation layer will be the public interface or API of the Catalogue. The main purpose of this layer is to provide its functionality to the Gadget Visual Storyboard, or any other third party IDE or application. The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs. REST is an architectural style, not a toolkit or a standard. Even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types. Another characteristic of the REST style is its stateless assumption. The Catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response. The API provided is explained in detail in Section~\ref{sec:catalogue_api}.

As part of the Presentation layer, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification. The SPARQL endpoint is mostly offered to enable other developers to query directly the Catalogue knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

\subsection{Core layer}

The Core or Business Logic layer contains all the FAST domain-specific processing. The resources and relations between each other, which define the domain model, are extracted from the FAST ontology~\cite{moeller2011fast_ontology}. These layer provides support to the functions defined in the API, and interact with the persistence layer to retrieve and store any information of the data model.

@@

At the bottom part of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services. These are RESTful services, SOAP-based web services, or similar. In FAST, these web services are wrapped, extracting their definition and encapsulation their behaviour. To creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service. However, the Catalogue does not crawl the Web looking for Web services. Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe\footnote{http://iserve.kmi.open.ac.uk/}, a platform for publishing Semantic Web Services as linked data, no matter their original format. Therefore, thousands of web services defined or annotated in a semantic way, are also served from the Catalogue, in a uniform manner.

The Catalogue is a semantic application. As a result, it has to handle with ontologies. Apart from the core ontologies being used across the whole application (see~\cite{moeller2011fast_ontology}), the building blocks may reference classes and properties from diverse vocabularies and ontologies. To have a better understanding of them, the Catalogue imports them inside its repository any time it detects a building block is using an unknown class, property, or in other words, an unknown ontology. While importing an ontology, the Catalogue tries to get its definition from the URI of a given class or property (as stated in~\cite{berrueta2008}). However, sometimes this may be not succesful, and the Catalogue will try to get its definition querying Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which cralws the web and keeps provenance metadata for the ontologies.

\subsection{Plumbing layer}

The Plumbing layer abstract the interaction with triple stores, RDF models and physical persistence to disk, database, etc. It provides the different mechanisms to allow the Catalogue the storage of the domain objects into a triple store in its underlying representation of RDF.

There are many triple stores solutions, any of them with its own advantages and disadvantages. To make the Catalogue independent of the solution adopted, it makes uses of RDF2Go, an abstraction over triple (and quad) stores. The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.

@@

The format supported, and in which they must be constructed, by every HTTP request is JSON\footnote{http://www.json.org/}. JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.

@@

\singlespacing
\begin{verbatim}
{
  "canvas": [
    { "uri": "http://catalogueURL/screens/238" },
    { "uri": "http://catalogueURL/screens/323" }
  ],
  "palette": [
    { "uri": "http://catalogueURL/screens/636" }
  ],
  "domainContext": {
    "tags": [
      {
        "label": { "en-GB": "Amazon" },
        "means": "http://dbpedia.org/page/Amazon.com"
      }
    ],
    "user": "http://ismaelrivera.es/#me",
  },
  "criterion": "reachability"
}
\end{verbatim}
\doublespacing

@@

\singlespacing
\begin{verbatim}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
    xmlns:foaf="http://xmlns.com/foaf/0.1/"
    xmlns:dc="http://purl.org/dc/terms/"
    xmlns:fgo="http://purl.oclc.org/fast/ontology/gadget#"
    xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
    xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#"
    xmlns:ctag="http://commontag.org/ns#">
  <rdf:Description rdf:about="http://demo.fast.morfeo-project.org/FASTCatalogue/screens/1">
    <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
    <rdfs:label xml:lang="en-gb">Product Search</rdfs:label>
    <rdfs:label xml:lang="en-es">Búsqueda de productos</rdfs:label>
    <rdfs:label xml:lang="en-de">Produktsuche</rdfs:label>
    <dc:description xml:lang="en-gb">On this screen, the user can enter a search
    term.</dc:description>
    <dc:description xml:lang="en-gb">En esta ventana, el usuario puede introducir
    el criterio de búsqueda.</dc:description>
    <dc:description xml:lang="en-gb">Auf diesem Screen kann der Benutzer einen
    Suchbegriff eingeben.</dc:description>
  </rdf:Description>
</rdf:RDF>
\end{verbatim}
\doublespacing

@@

In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood. The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it. In fact, each such representation has its own URI, under which it can be retrieved. Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}. In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \verb|http://catalogueURL/screens/235|), as well as the required representation format (e.g., JSON). The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \verb|http://catalogueURL/screens/235.json|), as illustrated in Fig.~\ref{fig:content_negotiation}.
On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.

@@

In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa. Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI. The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.

@@

\subsection{API Calls} % (fold)
\label{sub:api_calls}

@@

\subsection{Managing Building Blocks} % (fold)
\label{sub:managing_building_blocks}

Any building block within the Catalogue is a resource in terms of REST philosophy, therefore it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation. Two concepts should be clarified: \emph{collection}, as a set of resources which access URI is \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g. screens, operators), and \emph{member} of the collection, in other words, the building block itself, which access URI is \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block. The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}. These methods are to what is called \emph{Prototype} of building blocks (see \cite{moeller2011fast_ontology}).

\begin{table}[htb]
\caption{CRUD operations}
\label{tab:crud_operations}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
                         \textcolor{white}{HTTP method} &
                         \textcolor{white}{HTTP body} &
                         \textcolor{white}{Description}\\ \hline
Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
Collection URI & PUT & N/A & Not used.\\ \hline
Collection URI & POST & JSON representation of the building block & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection. The URI created is returned by this operation.\\ \hline
Collection URI & DELETE & N/A & Not used.\\ \hline
Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
Member URI & PUT & JSON representation of the building block & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
Member URI & POST & N/A & Not used. \\ \hline
Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
\end{tabular}
\end{center}
\end{table}

The JSON structure for create any building bock can be found in the appendixes. All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type. Screen-flows are defined in Appendix C, screens in Appendix D and forms, operators and back-end services share the same structure detailed in Appendix E. However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.

@@

\singlespacing
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "creator": "http://ismaelrivera.es/#me",
  "description": {"en-gb": "Please fill the description..."},
  "homepage": "http://fast.morfeo-project.eu/",
  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
  "id": "28",
  "label": {"en-gb": "Prueba1"},
  "postconditions": [
      {
        "id": "item",
        "label": { "en-gb": "An item" },
        "pattern": "?item
                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                    http://fast.morfeo-project.org/ontologies/amazon#Item",
        "positive": "true"
      }
    ]
  ],
  "preconditions": [
    [
      {
        "id": "filter",
        "label": { "en-gb": "A search request" },
        "pattern": "?search
                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                    http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
        "positive": "true"
      }
    ]
  ],
  "rights": "http://creativecommons.org/",
  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
  "tags": [{"label": {"en-gb": "amazon"}}],
  "type": "screen",
  "uri": "http://catalogueURL/screens/28",
  "version": "0.1"
}
\end{verbatim}
\doublespacing

@@

\subsubsection{GetMetadata} % (fold)

Retrieves all the metadata about a list of building blocks or classes (concepts) stored in the catalogue. All building blocks are supported, no matter if it is a prototype or a clone.

The URL to access this operation is \verb|http://catalogueURL/getmetadata| using HTTP POST method. Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters of a response.

\begin{table}[htb]
\caption{GetMetadata Request Parameters}
\label{tab:getmetadata_request}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
N/A & A list of URIs. & Required\\ \hline
\end{tabular}
\end{center}
\end{table}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\usepackage{amsmath}

@@

\begin{small}
\begin{tabular}{|l|l|l|p{6.6cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & Release for M24 \\
 & & Ismael Rivera (NUIG) & \\ \hline
2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Release for M36 \\
 & & Ismael Rivera (NUIG) & \\ \hline
\end{tabular}
\end{small}

@@

This sections enumerates and gives a brief overview of the changes made in the deliverable with regards to the version 1.0 released for the milestone M24.

\begin{itemize}
 \item The installation guide has been modified to include some minor changes in the configuration, and a better example for the ``Hello World!''.
 \item Two operations have been removed: 'Find' and 'Check' for screens, used at screen-flow design, since the same functionaly was offered by the operation Find\&Check, so no functionality is duplicated or overlapped. However, using this unique operation does not mean that every request does a \emph{find} and a \emph{check}, these actions are optional and can be specified in the request.
 \item There have been some substantial changes in the API methods Find\&Check and Planner and Metadata reflect the changes made in the Catalogue prototype.
 \item There have been created new operations in the API to support the prototype-based semantics for RDF as defined in D2.2.3~\cite{moeller2011fast_ontology}.
 \item The Catalogue Architecture has been updated to reflect the new functionality and other changes.
 \item There have been implemented new mechanisms for managing and importing ontologies. These mechanism are explained in~\ref{ssec:logic_tier}.
 \item A new appendix has been added in order to explain the decisions and logic behind the main methods of the API, how the building block recommendation works, algorithm supporting the building block planning, matching and inferencing in pre-/post-conditions, and the integration with iServe\footnote{http://iserve.kmi.open.ac.uk/}, enhancing the web service discovery.
 \item There are two new resources, and its CRUD operations: concepts (classes) and samples (instances).
\end{itemize}

@@

The following section covers the configuration options you may have to set up before start using the Catalogue. The configuration is done in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.

The resources (building blocks) created in the Catalogue have an unique identifier or URI. This URI is constructed relatively to the Catalogue URL which is configured by the parameter \emph{serverURL}. If you are testing the application in your machine, it may look like \verb|http://localhost:8080/FASTCatalogue|.

The FAST Catalogue relies on a Sesame repository for the persistent storage. Sesame repositories can be \emph{local} or \emph{remote}. Local repositories just need the parameter \emph{storageDir} pointing at a local directory, with write permission in order to allow Sesame to create and delete the necessary binary files to store and retrieve the data.

@@

Setting up a remote repository requires a few more steps. \emph{Remote} means the repository is installed in a web server and accessed via HTTP, independently if it is in the same machine or in a different dedicated machine. First the OpenRDF Sesame Server needs to be installed. The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}. Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier.

\begin{verbatim}
sesameServer=http://sesameServerURL/openrdf-sesame
repositoryID=repository-id
\end{verbatim}

Note: If you want to provide the data of in the repository via a SPARQL endpoint, the easiest way it is to configure the Catalogue using a remote repository, as in this case Sesame provides an HTTP accessible SPARQL endpoint by default.

@@

It is recommended to check if the FAST Catalogue has been successfully installed. The simpliest manner is to send a HTTP GET request to retrieve concepts (classes) and attributes (properties) of the predefined ontologies to \verb|http://catalogueURL/FASTCatalogue/concepts| (e.g. typing the address in a Web browser). The response is a list of all the concepts and their associated metadata, as in the following example:

\singlespacing
\begin{verbatim}
[
  {
    "attributes": [],
    "label": {"en": "A request to search for an item"},
    "tags": [{
      "label": {"en": "amazon"},
      "means": "http://www.amazon.com"
    }],
    "uri": "http://fast.morfeo-project.org/ontologies/amazon#SearchRequest"
  },
  {
    "attributes": [
      {
        "type": "http://xmlns.com/foaf/0.1/Document",
        "uri": "http://fast.morfeo-project.org/ontologies/amazon#smallImage"
      },
      {
        "type": "http://www.w3.org/2001/XMLSchema#double",
        "uri": "http://fast.morfeo-project.org/ontologies/amazon#price"
      }
    ],
    "label": {"en": "An item handled by the Amazon Service"},
    "tags": [{
      "label": {"en": "amazon"},
      "means": "http://www.amazon.com"
    }],
    "uri": "http://fast.morfeo-project.org/ontologies/amazon#Item"
  },
  ...
]
\end{verbatim}
\doublespacing

By default, the Catalogue does not include any building block, therefore any GET request for any type of building block, as for instance to \verb|http://catalogueURL/FASTCatalogue/screens| should return an empty list \verb|[]| (if JSON requested) or \verb|No screens found| (if HTML requested).

@@

This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or APIs is done. The Catalogue platform has a modular design, which are spread in basically three layers, depending on the nature of each component:
\begin{itemize}
 \item \emph{Presentation tier} is the top layer which represents the public API for the Catalogue. It brings together all the relaying functionality in a simple but flexible manner to be used by other IDEs or applications. In FAST, this is the unifying interface with other components such as the Gadget Visual Storyboard or the Service Wrapper.
 \item \emph{Logic tier} is the core layer which implements the business logic for the different processes involved in the Catalogue, and communicates with third party libraries or APIs.
 \item \emph{Data tier} is the bottom layer. It offers storage capabilities in triple stores, relational databases and binary files, and acts as an abstraction to RDF model logic.
\end{itemize}

\begin{figure}[htb]
\label{fig:catalogue_architecture}
\begin{center}
	\includegraphics[width=13cm]{images/catalogue_architecture}
	\caption{Catalogue's Architecture}
\end{center}
\end{figure}

\subsection{Presentation tier / Public API}

The Presentation tier is the public interface or API of the Catalogue. The main purpose of this layer is to provide its functionality to the Gadget Visual Storyboard, or any other third party IDE or application. The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs. REST is an architectural style, not a toolkit or a standard , even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types. Another characteristic of the REST style is its stateless assumption. The Catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response. The API provided is explained in detail in Section~\ref{sec:catalogue_api}.

As part of the Presentation tier, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification. The SPARQL endpoint is mostly offered to enable other developers to query directly the Catalogue knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

\subsection{Logic tier}
\label{ssec:logic_tier}

The Logic tier or business logic layer contains all the FAST domain-specific processing. The resources and relations between each other, which define the domain model, are extracted from the FAST ontology~\cite{moeller2011fast_ontology}. This layer provides support to the functions defined in the API, and interact with the data tier to retrieve and store any information of the data model.

@@

At the bottom level of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services. These are RESTful services, SOAP-based web services, or similar. In FAST, these web services are wrapped, extracting their definition and encapsulating their behaviour. The creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service. However, the Catalogue does not crawl the Web looking for Web services. Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe, a platform for publishing Semantic Web Services as linked data, no matter their original format. Therefore, thousands of web services defined or annotated in a semantic way, are also served from the Catalogue, in a uniform manner.

The Catalogue is a semantic application. As a consequence, it must handle and work with ontologies and vocabularies. The business model itself is described by in the FAST ontology, and the Catalogue goes along with a few established ontologies integrated in its core (see~\cite{moeller2011fast_ontology}). However, the definition of the pre-/post-conditions of the building blocks are open to any ontology or vocabulary, and they may reference classes and properties from any of them. To have a better understanding of these classes, properties, and so on, the Catalogue needs to import into its own knowledge base the ontology in which they are described along with the relations with other classes and properties. In order to adquire these vocabularies, it follows the basic recipe dictated in~\cite{berrueta2008} which says that deferencing a vocabulary URI should provide a machine-processable (RDF) content back. Nonetheless, this is just a good practice and not every ontology published on the Internet follows it. Hence, going a step further, whenever an ontology cannot be retrieved by deferencing its URI, the Catalogue queries Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which crawls and contains thousands of vocabularies and metadata as their provenance, which the Catalogue can take advantage of.

\subsection{Data tier}

The Data tier abstracts the interaction with triple stores, RDF models and physical persistence to disk, database, etc. It provides different mechanisms to allow the Catalogue the storage of domain objects into a triple store in its underlying representation of RDF.

There are many triple stores solutions, all of them with their own advantages and disadvantages. To make the Catalogue independent of the solution adopted, it makes uses of RDF2Go, an abstraction over triple (and quad) stores. The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.

@@

The body of the requests to the API must be in JSON\footnote{http://www.json.org/}. JSON is a lightweight data interchange format which simplicity has resulted in widespread use among web developers, easy to read and write and able to process and parse using any programming language because its structures map directly to data structures used in most programming languages.

@@

\singlespacing
\begin{verbatim}
{
  "canvas": [
    { "uri": "http://catalogueURL/screens/clones/2338" },
    { "uri": "http://catalogueURL/screens/clones/1323" }
  ],
  "palette": [
    { "uri": "http://catalogueURL/screens/636" }
  ],
  "domainContext": {
    "tags": [
      {
        "label": { "en-GB": "Amazon" },
        "means": "http://dbpedia.org/page/Amazon.com"
      }
    ],
    "user": "http://ismaelrivera.es/#me",
  },
  "criterion": "reachability"
}
\end{verbatim}
\doublespacing

@@

\singlespacing
\begin{verbatim}
<rdf:RDF
    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
    xmlns:foaf="http://xmlns.com/foaf/0.1/"
    xmlns:dc="http://purl.org/dc/terms/"
    xmlns:fgo="http://purl.oclc.org/fast/ontology/gadget#"
    xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
    xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#"
    xmlns:ctag="http://commontag.org/ns#">
  <rdf:Description rdf:about="http://catalogueURL/screens/1">
    <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
    <rdfs:label xml:lang="en-gb">Product Search</rdfs:label>
    <rdfs:label xml:lang="en-es">Búsqueda de productos</rdfs:label>
    <rdfs:label xml:lang="en-de">Produktsuche</rdfs:label>
    <dc:description xml:lang="en-gb">On this screen, the user can enter a search
    term.</dc:description>
    <dc:description xml:lang="en-gb">En esta ventana, el usuario puede introducir
    el criterio de búsqueda.</dc:description>
    <dc:description xml:lang="en-gb">Auf diesem Screen kann der Benutzer einen
    Suchbegriff eingeben.</dc:description>
  </rdf:Description>
</rdf:RDF>
\end{verbatim}
\doublespacing

@@

In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood. The URI of a particular building block must be understood as the identifier of the building block such, as opposed to a particular representation (JSON, HTML, etc.) of it. In fact, each such representation has its own URI, under which it can be retrieved. Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}. In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \verb|http://catalogueURL/screens/235|), as well as the required representation format (e.g., JSON). The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \verb|http://catalogueURL/screens/235.json|), as illustrated in Fig.~\ref{fig:content_negotiation}.
On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations is most appropriate for a given request.

@@

In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa. Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI. The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.ttl}, \texttt{.json} or \texttt{.html} for RDF/XML, Turle, JSON or HTML respectively.

@@

\subsection{API calls} % (fold)
\label{sub:api_calls}

@@

\subsection{Managing building blocks} % (fold)
\label{sub:managing_building_blocks}

Any building block within the Catalogue is a resource in terms of REST philosophy, therefore it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation. Two concepts should be clarified: \emph{collection}, as a set of resources are accessed at the URL \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g. screens, operators), and \emph{member} of the collection, in other words, the building block itself, being accessed at the URL \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block. The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.

\begin{table}[htb!]
\caption{CRUD operations}
\label{tab:crud_operations}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
                         \textcolor{white}{HTTP method} &
                         \textcolor{white}{HTTP body} &
                         \textcolor{white}{Description}\\ \hline
Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
Collection URI & PUT & N/A & Not used.\\ \hline
Collection URI & POST & JSON representation of the building block & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection. The URI created is returned by this operation.\\ \hline
Collection URI & DELETE & N/A & Not used.\\ \hline
Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
Member URI & PUT & JSON representation of the building block & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
Member URI & POST & N/A & Not used. \\ \hline
Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
\end{tabular}
\end{center}
\end{table}

The JSON structure for creating any building block can be found in the appendixes. All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type. Screen-flows are defined in Appendix C, screens in Appendix D and forms, operators and back-end services share the same structure detailed in Appendix E. However, a few examples of request and responses are shown in order to clarify how the requests are constructed and sent, and how the responses look like.

@@

\singlespacing
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "creator": "http://ismaelrivera.es/#me",
  "description": {"en-gb": "Please fill the description..."},
  "homepage": "http://fast.morfeo-project.eu/",
  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
  "id": "28",
  "label": {"en-gb": "Prueba1"},
  "postconditions": [
      {
        "id": "item",
        "label": { "en-gb": "An item" },
        "pattern": "?item
                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                    http://fast.morfeo-project.org/ontologies/amazon#Item",
        "positive": "true"
      }
    ]
  ],
  "preconditions": [
    [
      {
        "id": "filter",
        "label": { "en-gb": "A search request" },
        "pattern": "?search
                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                    http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
        "positive": "true"
      }
    ]
  ],
  "rights": "http://creativecommons.org/",
  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
  "tags": [{"label": {"en-gb": "amazon"}}],
  "uri": "http://catalogueURL/screens/28",
  "version": "0.1"
}
\end{verbatim}
\doublespacing

@@

\subsubsection{Metadata} % (fold)
\label{ssub:metadata}

Retrieves all the metadata about a list of building blocks or classes (concepts) stored in the catalogue. This functionality is already provided by~\ref{sub:managing_building_blocks} if one specific building block, or the entire collection, is to be retrieved. However, this operation provides a way to get the metadata of an arbitrary number of resources, specified by their URIs, with no disctintion of their type.

The URL to access this operation is \verb|http://catalogueURL/metadata|. The operation is invoked sending a HTTP POST request, Table~\ref{tab:metadata_request} shows the parameters needed for the invocation and Table~\ref{tab:metadata_response} details the different parameters of the response.

\begin{table}[htb!]
\caption{Metadata Request Parameters}
\label{tab:metadata_request}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
N/A & A list of URIs. & Required\\ \hline
\end{tabular}
\end{center}
\end{table}
